\chapter*{Abstract}
\pagestyle{fancy}

As the decentralized finance (DeFi) landscape continues to evolve, slippage and Maximal Extractable Value (MEV) remain significant hurdles for users on Ethereum and EVM-compatible chains. While CoW DEX solutions have successfully mitigated these issues in the EVM space, the Sui network presents a unique environment where its Object-Centric model inherently provides robust protection against certain forms of MEV. Nevertheless, slippage continues to be a persistent challenge for decentralized exchanges (DEXs) on the Sui network.

\noindent This thesis focuses on the analysis, architectural design, and development of a CoW DEX on Sui, leveraging the network's object-oriented paradigm and inspired by the core architecture of the Ethereum CoW Protocol. The research proposes a novel auction mechanism incorporating a Commit-Reveal phase and utilizes the Hot Potato Pattern to ensure atomicity, fairness, and the prevention of solver fraud. The system aims to establish an intent-based DEX aggregator that delivers optimal pricing and zero slippage for coincident orders (Coincidence of Wants), while effectively harnessing Sui's core liquidity layers, specifically Deepbook. This work contributes a scalable and secure trading framework that enhances transparency and capital efficiency within the Sui ecosystem.

\vspace{1cm}

\keywords{Blockchain, Decentralized Finance, Coincidence of Wants, Slippage, Maximal Extractable Value, Intent-based, Object-Centric, Sui, Ethereum, Deepbook, CoW Protocol.}

\vspace*{\fill}
\newpage
